---
title: "PINP Code for seabird detection in fur seal diets"
author: "Natasha Hardy"
date: "25/02/2021"
output: html_document
---

## About

**Citation:** MS in prep. Add citation after submission.

This document contains all code pertinent to figures and analyses presented in this publication:
(i) data manipulation, 
(ii) data analyses and 
(iii) visualisations for the seabird detection in long-nosed fur seal diet project.

Predator species: long-nosed fur seals.
Prey target taxa: seabirds.
Total samples: 99 scats from long-nosed fur seals.
Locations: Cape Bridgewater, Barunguba (Montague Island), Gabo Island and Deen Maar (Lady Julia Percy).
Sample collection dates: Jan 2015 - Jan 2017.
Project objective: to compare the detection rates of birds in long-nose fur seal scats across the range of this predator in SE Australia from Bass Strait to southern NSW, and over multiple seasons.

## Workspace

```{r Workspace, echo=TRUE, message=FALSE, results='hide'}

#Check working directory - the same as the doc directory.

#getwd()

#Install Packages as needed
#install.packages("rcompanion")

#Load
library(tidyverse)
library(tidyr)
library(tidytext)
#library(reshape)
library(plyr)
library(dplyr)
library(PNWColors)
library(lattice)
library(reshape2)
library(lme4)
library(ggplot2)
library(rcompanion)
library(gg.gap)
library(forcats)
library(devtools)
library(cowplot)
library(grid)
library(gridExtra)
library(scales)
library(ggpubr)
library(here)
"%notin%" = Negate('%in%')
here::here()

#Note that loading all these packages is hell because of overlap in functions, some packages being superceded but still requiring each other or requiring functions from one package to another. You'll notice I had to force a lot of functions by calling the package as follows 
#data %>% package::function(etc). I'm sure there are more talented and savvy coders than me who can figure this out more elegantly.

```

## Load data

These dataframes are currently downloaded as .csv spreadsheets from the diet analysis master document found here: 
[PINP Bird Data](https://docs.google.com/spreadsheets/d/1phfyo3kUyWZfwKF0IAHXgkxy23dbYs_tKkCfEIWeY8g/edit#gid=2115650924)

**NOTE:** we want to clean this up and include the whole spreadsheet and read these in as tabs.

```{r Load data}

#0 = "hp_taxa" tab
hp_taxa_export <- read.csv(here("data/input_data/hp_taxa.csv"), header=TRUE)

#1 = "bird_detect_med" tab
bird_wide_med <- read.csv(here("data/input_data/bird_detect_med.csv"), header=TRUE)

#2 = "bird_detect_high" tab
bird_wide_high <- read.csv(here("data/input_data/bird_detect_high.csv"), header=TRUE)

#3 = "penguin_detect_med" tab
penguin_wide_med <- read.csv(here("data/input_data/penguin_detect_med.csv"), header=TRUE)

#4 = "penguin_detect_high" tab
penguin_wide_high <- read.csv(here("data/input_data/penguin_detect_high.csv"), header=TRUE)

#5 = "otu_med" tab
otu_med_export <- read.csv(here("data/input_data/otu_med.csv"), header=TRUE)

#6 = "otu_high" tab
otu_high_export <- read.csv(here("data/input_data/otu_high.csv"), header=TRUE)

#7 penguin haplotypes
penguin_haplos_export <- read.csv(here("data/input_data/penguin_haplos.csv"), header=TRUE)

```


## Data manipulations

Here we mainly clean column or variable names, gather wide data and change to longform.
I originally chose to input data as columns in the master document tab "BIRD_DATA" in PINP_BIRD_DATA.xlsx on GoogleDrive. This is because many columns were calculated based off other columns (such as medium stringency DNA OTU total abundance used to calculate a 1% cut-off for high stringency DNA OTU columns). This means a little data wrangling on this end to merge these data into long-form for analyses and graphing.

For the following manipulations, I'm going to:
(i) Rename colnames for each of the four datasets exported, 
(ii) Transform from wide to long data, using melt(), and include arguments for renaming the output variable and value to "metric" and "detection", and
(iii) Add a categorical descriptor for each method, to later merge datasets 1+2 (bird detections) and 3+4 (penguin detections)

Columns are renamed as follows:
hp = hard-part analysis to ID prey
dna = DNA metabarcoding to ID prey
dna_or_hp = detections using either method
same_taxon = both detected the same bird taxon (for bird only, as "both" and "same" are equal for looking at one taxon such as penguins)
Note, I have removed "both_methods = both detected birds/penguins". This column was confusing and did not add information, it was originally to report the number of samples that contained different taxa identified from the same taxa using both methods.

Then add descriptor:
medium = standard QC for DNA based analysis
high = additional exclusion of trace amounts of DNA that passed QC

Note on 22nd March 2020, I corrected a mistake in an excel function for the penguin datasets, that had effectively inputted some wrong numbers in the penguin detection columns. I have rerun analyses and updated stats, and figures.

```{r Data 0: Hard part data}

hp_taxa = hp_taxa_export %>%
  dplyr::rename(`E. minor` = `SUM.of.penguin_hp`, 
                Procellarid = `SUM.of.procellarid_hp`,
                `M. serrator` = `SUM.of.gannet_hp`, 
                `Seabird unknown` = `SUM.of.unknown_hp`) %>%
  #Renamed columns
  reshape2::melt(id.vars=c("sample_id", "location_id", "location", "time_id", "group_id"), variable.name = "Taxa", value.name = "detect") %>%
  #Transform wide to long data, using melt(), and added arguments for renaming the output variable and value to "metric" and "detection"
  #Added the character vector/factor for merging with the high stringency data
  mutate(method = "hard-parts")

#Rename factor levels
hp_taxa$location <- factor(hp_taxa$location, 
                           levels = c("Cape Bridgewater", "Gabo Island", "Lady Julia Percy", "Montague Island"), 
                           labels = c("Cape Bridgewater", "Gabo Island", "Deen Maar", "Barunguba"))

hp_taxa$time_id <- factor(hp_taxa$time_id, 
                          levels = c("Sep_15", "Jan_16", "Sep_16", "Jan_17", "Mar_17"), 
                          labels = c("Sep'15", "Jan'16", "Sep'16", "Jan'17", "Mar'17"))

```

```{r Data 1: Bird standard QC}

bird_med = bird_wide_med %>% 
  dplyr::rename(hp = `SUM.of.allbirds_hp`, dna = `SUM.of.allbirds_dna_cons`,
                dna_or_hp = `SUM.of.allbirds_cons_dna_or_hp`, 
                both_methods = `SUM.of.allbirds_cons_both`, 
                same_taxon = `SUM.of.allbirds_cons_sametaxa`) %>%
  #Renamed columns
  dplyr::select(., -both_methods) %>%
  reshape2::melt(id.vars=c("sample_id", "location_id", "location", "time_id", "group_id"), variable.name = "metric", value.name = "detection") %>%
  #Transform wide to long data, using melt(), and added arguments for renaming the output variable and value to "metric" and "detection"
  #Added the character vector/factor for merging with the high stringency data
 mutate(method = if_else(metric == "hp", "hard part", "DNA standard QC"))

#Rename factor levels
bird_med$location <- factor(bird_med$location, 
                            levels = c("Cape Bridgewater", "Gabo Island", "Lady Julia Percy", "Montague Island"), 
                            labels = c("Cape Bridgewater", "Gabo Island", "Deen Maar", "Barunguba"))

bird_med$time_id <- factor(bird_med$time_id, 
                           levels = c("Sep_15", "Jan_16", "Sep_16", "Jan_17", "Mar_17"), 
                           labels = c("Sep'15", "Jan'16", "Sep'16", "Jan'17", "Mar'17"))

```

```{r Data 2: Bird high QC}

bird_high = bird_wide_high %>% 
  dplyr::rename(hp = `SUM.of.allbirds_hp`, dna = `SUM.of.allbirds_dna_high`, 
                dna_or_hp = `SUM.of.allbirds_high_dna_or_hp`, 
                both_methods = `SUM.of.allbirds_high_both`, 
                same_taxon = `SUM.of.allbirds_high_same`) %>%
  #Renamed columns
  dplyr::select(., -both_methods) %>%
  reshape2::melt(id.vars=c("sample_id", "location_id", "location", "time_id", "group_id"), variable.name = "metric", value.name = "detection") %>%
  #Transform wide to long data, using melt(), and added arguments for renaming the output variable and value to "metric" and "detection"
  filter(metric != "hp") %>% #here I'm removing the repeated subset of data on hard-part analysis for graphing
  mutate(method = "DNA stringent QC")

#Rename factor levels
bird_high$location <- factor(bird_high$location, 
                             levels = c("Cape Bridgewater", "Gabo Island", "Lady Julia Percy", "Montague Island"), 
                             labels = c("Cape Bridgewater", "Gabo Island", "Deen Maar", "Barunguba"))

bird_high$time_id <- factor(bird_high$time_id, 
                            levels = c("Sep_15", "Jan_16", "Sep_16", "Jan_17", "Mar_17"),
                            labels = c("Sep'15", "Jan'16", "Sep'16", "Jan'17", "Mar'17"))

```

```{r Data 3: Penguin standard QC}

penguin_med = penguin_wide_med %>% 
  dplyr::rename(hp = `SUM.of.penguin_hp`, dna = `SUM.of.penguin_dna_cons`,
                dna_or_hp = `SUM.of.penguin_cons_dna_or_hp`, 
                same_taxon = `SUM.of.penguin_cons_both`) %>%
  #Renamed columns
  reshape2::melt(id.vars=c("sample_id", "location_id", "location", "time_id", "group_id"), variable.name = "metric", value.name = "detection") %>%
  #Transform wide to long data, using melt(), and added arguments for renaming the output variable and value to "metric" and "detection"
   mutate(method = if_else(metric == "hp", "hard part", "DNA standard QC"))

#Rename factor levels
penguin_med$location <- factor(penguin_med$location, 
                               levels = c("Cape Bridgewater", "Gabo Island", "Lady Julia Percy", "Montague Island"), 
                               labels = c("Cape Bridgewater", "Gabo Island", "Deen Maar", "Barunguba"))

penguin_med$time_id <- factor(penguin_med$time_id, 
                              levels = c("Sep_15", "Jan_16", "Sep_16", "Jan_17", "Mar_17"), 
                              labels = c("Sep'15", "Jan'16", "Sep'16", "Jan'17", "Mar'17"))

```

```{r Data 4: Penguin high QC}

penguin_high = penguin_wide_high %>% 
  dplyr::rename(hp = `SUM.of.penguin_hp`, dna = `SUM.of.penguin_dna_high`,
                dna_or_hp = `SUM.of.penguin_high_dna_or_hp`, 
                same_taxon = `SUM.of.penguin_high_both`) %>%
  #Renamed columns
  reshape2::melt(id.vars=c("sample_id", "location_id", "location", "time_id", "group_id"), variable.name = "metric", value.name = "detection") %>%
  #Transform wide to long data, using melt(), and added arguments for renaming the output variable and value to "metric" and "detection"
  #Added the character vector/factor for merging with the high stringency data
  filter(metric != "hp") %>% #here I'm removing the repeated subset of data on hard-part analysis for graphing
  mutate(method = "DNA stringent QC")

#Rename factor levels
penguin_high$location <- factor(penguin_high$location, 
                                levels = c("Cape Bridgewater", "Gabo Island", "Lady Julia Percy", "Montague Island"), 
                                labels = c("Cape Bridgewater", "Gabo Island", "Deen Maar", "Barunguba"))

penguin_high$time_id <- factor(penguin_high$time_id, 
                               levels = c("Sep_15", "Jan_16", "Sep_16", "Jan_17", "Mar_17"), 
                               labels = c("Sep'15", "Jan'16", "Sep'16", "Jan'17", "Mar'17"))

```

```{r Data 5: OTU standard QC}

#OTU_med cleaning

otu_med = otu_med_export %>% 
  dplyr::rename(OTU1_E_minor = `SUM.of.cons_OTU_1..Eudyptula.minor.`, 
                OTU2_Procellarid1 = `SUM.of.cons_OTU_2..Procellariidae.sp.1.`, 
                OTU3_Procellarid2 = `SUM.of.cons_OTU_3..Procellariidae.sp.2.`, 
                OTU4_T_melanophris = `SUM.of.cons_OTU_4..Thalassarche.melanophris.`,
                OTU5_S_bergii = `SUM.of.cons_OTU_5..Sterna.bergii.`) %>%
  #Renamed columns
  reshape2::melt(id.vars=c("sample_id", "location_id", "location", "time_id", "group_id"), variable.name = "OTU", value.name = "seq_abun") %>%
  #Transform wide to long data, using melt(), and added arguments for renaming the output variable and value to "metric" and "detection"
  #Added the character vector/factor for merging with the high stringency data
  mutate(method = "DNA standard QC")

#Relative read abundances by taxon / OTU
#Truncating datasets to calculate additional sequence metrics.
mdata1 = otu_med %>% 
  filter(OTU == "OTU1_E_minor") %>% 
  mutate(seq_prop = seq_abun/sum(seq_abun)*100)
mdata2 = otu_med %>% 
  filter(OTU == "OTU2_Procellarid1") %>% 
  mutate(seq_prop = seq_abun/sum(seq_abun)*100)
mdata3 = otu_med %>% 
  filter(OTU == "OTU3_Procellarid2") %>% 
  mutate(seq_prop = seq_abun/sum(seq_abun)*100)
mdata4 = otu_med %>% 
  filter(OTU == "OTU4_T_melanophris") %>% 
  mutate(seq_prop = seq_abun/sum(seq_abun)*100)
mdata5 = otu_med %>% 
  filter(OTU == "OTU5_S_bergii") %>% 
  mutate(seq_prop = seq_abun/sum(seq_abun)*100)

#Join these subsetted and manipulated datasets
otu_med_data = rbind(mdata1, mdata2, mdata3, mdata4, mdata5) 

#These two are split due to problems knitting the doc
otu_med_data = otu_med_data %>%
  mutate(trace_high = if_else(otu_med_data$seq_abun >= 10, "high", "trace"),
         total_prop = seq_abun/sum(seq_abun)*100) #Add trace vs. high DNA abundance, and proportion of total seq abundance variable #Relative read abundance / total DNA abundance

#Change labels for location
otu_med_data$location <- factor(otu_med_data$location, levels = c("Cape Bridgewater", "Gabo Island", "Lady Julia Percy", "Montague Island"), labels = c("Cape Bridgewater", "GI", "DM", "Barunguba"))
#Change labels for time
otu_med_data$time_id <- factor(otu_med_data$time_id, levels = c("Sep_15", "Jan_16", "Sep_16", "Jan_17", "Mar_17"), labels = c("Sep'15", "Jan'16", "Sep'16", "Jan'17", "Mar'17"))
#Change labels for OTUs
#levels(otu_med_abundant$OTU)
otu_med_data$OTU <- factor(otu_med_data$OTU, levels = c("OTU1_E_minor", "OTU2_Procellarid1", "OTU3_Procellarid2", "OTU4_T_melanophris", "OTU5_S_bergii"), labels = c("E. minor", "Procellarid sp1", "Procellarid sp2", "T. melanophris", "S. bergii"))

otu_med_data$trace_high <- as.factor(otu_med_data$trace_high)

#output = otu_med_data

```

```{r Data 6: OTU RRA + Total Abun}

# Created a separate summary df
otu_med_sum <- ddply(otu_med_data, c("location", "time_id", "group_id", "sample_id"), summarise,
                  sample_abun = sum(seq_abun))
# Joined back by sample_id (shared variable)
otu_sampleabun <- merge(x = otu_med_data, y = otu_med_sum, by = c("location", "time_id", "group_id", "sample_id"))

otu_metrics = otu_sampleabun %>%
  select(location:OTU, method, trace_high, seq_abun, sample_abun, seq_prop, total_prop) %>%
  #dplyr::rename(location = `location.x`, time_id = `time_id.x`, group_id = `group_id.x`) %>%
  mutate(prop_abun = if_else(sample_abun >= 1, seq_abun/sample_abun*100, 0))

#output = otu_metrics and supercedes otu_med_data

```

### OTU metric outputs

Note the following output columns:
*seq_abun* = absolute sequence abundance for a given taxa within the sample;
*sample_abun* = total abundance of sequences obtained within that sample from any taxon;
*seq_prop* = proportional sequence abundance for a certain taxon within a sample, as a percent of total sequences belonging to that taxon;
*total_prop* = proportional sequence abundance for a certain taxon within a sample as a percent of total sequence abundance;
*prop_abun* = proportional sequence abundance for a certain taxon as a percent of total sequences within a sample;

```{r Data 7: Haplotypes}

penguin_haplos = penguin_haplos_export %>% 
  dplyr::select(-time) %>%
  dplyr::rename(haplo_1 = `SUM.of.haplo_1`, 
                haplo_2 = `SUM.of.haplo_2`, 
                haplo_3 = `SUM.of.haplo_3`,
                haplo_4 = `SUM.of.haplo_4`,
                haplo_5 = `SUM.of.haplo_5`,
                haplo_6 = `SUM.of.haplo_6`,
                haplo_7 = `SUM.of.haplo_7`) %>%
  #Renamed columns
  reshape2::melt(id.vars=c("sample_id", "location_id", "location", "time_id", "group_id"), variable.name = "haplotype", value.name = "seq_abun") %>%
  mutate(seq_presence = if_else(seq_abun > 0, "1", "0")) %>%
  filter(seq_abun > 1) 

penguin_haplos_sample = penguin_haplos %>%
  separate(sample_id, c("loc", "month", "year", "sample"))

penguin_haplos_sample$sample <- as.factor(penguin_haplos_sample$sample)

#output = penguin_haplos_sample

```

## Data outputs - masters & summaries

```{r Bird master}

#Need to join the bird_med and bird_high datasets for graphing

bird_master <- rbind(bird_med, bird_high)

```

```{r Penguin master}

#Need to join the penguin_med and penguin_high datasets for graphing

penguin_master <- rbind(penguin_med, penguin_high)

```

```{r Data summaries by metric}

#Overall bird detection by method: need to summarise of all samples containing bird by method of detection
#Need to scale to 100%

# Bird master
bird_sum_master <- ddply(bird_master, c("method", "metric"), summarise,
                  `Detected` = sum(detection, na.rm=TRUE)*100/99)

# Penguin master
penguin_sum_master <- ddply(penguin_master, c("method", "metric"), summarise,
                  `Detected` = sum(detection, na.rm=TRUE)*100/99)

#Can double check this with table()
#i.e., table(bird_long$metric, bird_long$value)

```


```{r Data summaries by location/time - total samples, message=FALSE, results='hide'}
#Overall detection by location and sampling time
#Here we want to select just the metrics that we want to display and need to scale to 100%

as.factor(bird_master$method)
levels(bird_master$method)

#0 sampling effort by time / location for Bec:
observations = bird_med %>% 
  group_by(location, time_id, group_id, metric) %>% 
  dplyr::summarize(obs = n()) %>%
  filter(metric == c('hp')) %>%
  select(location, time_id, group_id, obs)

write.csv(observations, here("data/output_data/sampling_effort.csv"))

```

```{r Data summaries by location/time - seabirds, message=FALSE, results='hide'}

#1 Bird detections by location/time
#Summarise
#PROBLEM = how to turn this into a percentage of samples within each location/time?

bird_sum_group = bird_med %>% 
  group_by(location, time_id, group_id, metric) %>% 
  dplyr::summarize(obs = n(), detected = sum(detection), percentage_detected = (detected/obs*100))

#Take subset for figure
bird_master_group = bird_sum_group %>%
  select(location, time_id, group_id, metric, obs, detected, percentage_detected) %>%
  filter(metric == c('hp', 'dna'))

```

```{r Data summaries by location/time - penguins, message=FALSE, results='hide'}

#2 Penguin detections by location/time

#Summarise

penguin_sum_group = penguin_med %>% 
  group_by(location, time_id, group_id, metric) %>% 
  dplyr::summarize(obs = n(), detected = sum(detection), percentage_detected = (detected/obs*100))

#Take subset for figure
penguin_master_group = penguin_sum_group %>%
  select(location, time_id, group_id, metric, obs, detected, percentage_detected) %>%
  filter(metric == c('hp', 'dna'))

```

```{r OTU & HP master}

#Need to join the penguin_med and penguin_high datasets for graphing

otu_hp_master = otu_metrics %>%
  dplyr::rename(`Taxa` = `OTU`) %>%
  mutate(detect = if_else(seq_abun > 0, 1, 0)) %>%
  full_join(hp_taxa, by= c("sample_id", "location_id", "location", "time_id", "group_id", "method", "Taxa", "detect")) %>%
  separate(sample_id, c("loc", "month", "year", "sample")) %>%
  select(sample, location_id, location, time_id, group_id, Taxa, method, trace_high, seq_abun, sample_abun, seq_prop, total_prop, prop_abun, detect)

#Change order of levels for Taxa

otu_hp_master$Taxa <- factor(otu_hp_master$Taxa, levels = c("E. minor", "Procellarid sp1", "Procellarid sp2", "Procellarid", "T. melanophris", "S. bergii", "M. serrator", "Seabird unknown"))
otu_hp_master$method <- as.factor(otu_hp_master$method)
otu_hp_master$sample <- as.factor(otu_hp_master$sample)
otu_hp_master$detect <- as.integer(otu_hp_master$detect)

```


## MS Figures

**Note on figures in MS:**

- Figure 1 was prepared by Harley at PINP using QGIS software. 

- Figure 2 is a composite panel of photos by Natasha Hardy & Rebecca McIntosh.


### Bar plots for MS:

**- Figure 3a** aims to display the proportion of samples detecting seabirds using either hard-part methods, DNA-based methods, both combined or either of those methods.

```{r Figure 3a Bird detections / metric}

#Figure 3a

#Figure 3 colour palette
fig3.pal <- pnw_palette(5, name = "Bay", type = "continuous")

#Relevel the method so that hard-part method comes first in the graph AND relabel
#levels(bird_sum_master$method) #check levels
bird_sum_master$method <- factor(bird_sum_master$method, levels = c("hard part", "DNA standard QC", "DNA stringent QC"), label = c("hard parts", "DNA (all)", "DNA (abundant)"))

#Relabel to clarify calculations
#levels(bird_sum_master$metric) # check levels
bird_sum_master$metric <- factor(bird_sum_master$metric, levels = c("hp", "dna", "dna_or_hp", "same_taxon"), label = c("hp", "dna", "dna or hp", "same taxon"))

#Figure 3a
fig3a_bird <- ggplot(bird_sum_master, aes(x=metric, y=Detected, fill=metric)) +
  geom_bar(stat="identity", position="dodge", width=1)+
  ggtitle("a) Seabird detections by metric")+
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        plot.title=element_text(size=18), #face="bold", 
        axis.title=element_blank(), #element_text(size=18)
        axis.text=element_text(size=18), 
        axis.text.x=element_blank(),
        legend.text=element_text(size=16), 
        legend.title=element_text(size=16), 
        legend.position="bottom"
        ) +
  ylim(0,50)+
  #theme(legend.position=c(0.60,0.80), legend.justification=c(0.4,0.5))+
  scale_fill_manual(values=fig3.pal)+
  guides(fill=FALSE) +
  #guides(fill=guide_legend(title="Containing:"))+
  #labs(x=NULL, y="Percentage of samples (%)") + #x="Metric", 
  facet_grid(.~method, scales="free",space='free') +
  theme(strip.text = element_text(size=rel(1.25)), #face="bold", 
        strip.background = element_rect(fill="lightgrey", colour="black",
                                        size=1))
fig3a_bird

#ggsave(here('figures/Final versions/fig3a_bird.jpeg'), plot = fig3a_bird, width = 9, height = 5, dpi = 300)

```

**- Figure 3b** aims to display the proportion of samples detecting penguins using either hard-part methods, DNA-based methods, both combined or either of those methods.

```{r Figure 3b Penguin detections / metric}

#Figure 3b

#Relevel the method so that hard-part method comes first in the graph
#levels(penguin_sum_master$method)
penguin_sum_master$method <- factor(penguin_sum_master$method, levels = c("hard part", "DNA standard QC", "DNA stringent QC"))

#Relabel to clarify
penguin_sum_master$method <- factor(penguin_sum_master$method, levels = c("hard part", "DNA standard QC", "DNA stringent QC"), label = c("hard parts", "DNA (all)", "DNA (abundant)"))

#levels(penguin_sum_master$metric)
penguin_sum_master$metric <- factor(penguin_sum_master$metric, levels = c("hp", "dna", "dna_or_hp", "same_taxon"), label = c("hp", "dna", "dna or hp", "same taxon"))

#Figure 3b
fig3b_penguin <- ggplot(penguin_sum_master, aes(x=metric, y=Detected, fill=metric)) +
  geom_bar(stat="identity", position="dodge", width=1)+
  ggtitle("b) Penguin detections by metric")+
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        plot.title=element_text(size=18), #face="bold", 
        axis.title=element_blank(), #element_text(size=18)
        axis.text=element_text(size=18), 
        axis.text.x=element_blank(),
        legend.text=element_text(size=16), 
        legend.title=element_text(size=16), 
        legend.position="bottom"
        ) +
  #theme(legend.position=c(0.60,0.80), legend.justification=c(0.4,0.5))+
  ylim(0,50)+
  scale_fill_manual(values=fig3.pal)+
  guides(fill=guide_legend(title="Containing:"))+
  labs(x=NULL, y="Percentage of samples (%)") + #x="Metric", 
  facet_grid(.~method, scales="free",space='free') +
  theme(strip.text = element_text(size=rel(1.25)), #face="bold", 
        strip.background = element_rect(fill="lightgrey", colour="black",
                                        size=1))
fig3b_penguin

#ggsave(here('figures/Final versions/fig3b_penguin.jpeg'), plot = fig3b_penguin, width = 9, height = 5, dpi = 300)


```

**- Figure 3** - merges panels a and b, and displays the overall detection rates for seabirds and penguins according the method of scat analysis.

```{r Figure 3 PANEL - summary stats}

#Merge the following plots into a single panel
#fig3a_bird & fig3b_penguin

#create common axis titles
y.grob <- textGrob("Percentage of samples (%)", gp=gpar(fontsize=18), rot=90) #fontface="bold", col="blue", 
#x.grob <- textGrob("Percentage of samples containing remains (%)", gp=gpar(fontsize=18)) #fontface="bold", col="blue", 

#combine using cowplot
fig3_panel = plot_grid(fig3a_bird, fig3b_penguin, 
                       #align = "h",
                       nrow = 2
                          #labels = c("a) Seabirds", "b) Little penguins (Eudyptula minor)"), 
                          #label_size = 18, ncol = 2, 
                          #label_x = 0.20, label_y = 1, hjust = 0, vjust = 1.5,
                          #scale = 0.95,
                          #label_x = 0.10, vjust = 1, rel_widths = c(1, 1), align = "v", 
                          #greedy = TRUE
                          ) #hjust = -7.5, label_x = c(0.12, 0.12, 0.12, 0.12, 0.12, 0.12)

#fig3_panel # check this

#add to plot
fig3_panel_ms = grid.arrange(arrangeGrob(fig3_panel, left = y.grob, padding = unit(1, "line")))
fig3_panel_ms

#ggsave(here("figures/Final versions/fig3_panel_ms.jpeg"), plot = fig3_panel_ms, width = 10, height = 10, scale = 0.95, dpi = 300)

#ggsave(here("figures/Final versions/fig3_panel_ms.pdf"), plot = fig3_panel_ms, width = 10, height = 10, scale = 0.95, dpi = 300)

```
